The universal property of direct integrals in W$^*$-categories

### 2018-06-29 direct_integrals.tex
I've added some stuff about compact measure spaces and a few corrections.

% categories
\newcommand{\cat}[1]{\mathsf{#1}}		% generic category
\newcommand{\Set}{\mathsf{Set}}
\newcommand{\Vect}[1]{\mathsf{Vect}_{#1}}	% vector spaces over a field
\newcommand{\Ban}{\mathsf{Ban}}			% (complex) Banach spaces
\newcommand{\Hilb}{\mathsf{Hilb}} 		% Hilb, cat of Hilbert spaces
\newcommand{\Rep}[1]{\mathsf{Rep}(#1)}		% cat normal Hilbert space reps of a C$^*$-algebra
\newcommand{\NRep}[1]{\mathsf{nRep}(#1)}	% cat of normal Hilbert space reps of a W$^*$-algebra
\newcommand{\id}[1]{1_{#1}}			% identity morphism

@@

We let $\LMeas$ be the category of localizable measurable spaces~\cite{nlab:localizable_measure}. So objects of $\LMeas$ are triples $(X,\Sigma,\Null)$, where $(X,\Sigma)$ is a $\sigma$-algebra and $\Null\subseteq\Sigma$ is a $\sigma$-ideal, whose elements are called \emph{null sets}, such that $\Sigma/\Null$ is complete as a Boolean algebra. Measurable sets with null complement are also called \emph{full sets}. We usually just write $X$ for a localizable measurable space, and indicate its $\sigma$-algebra and $\sigma$-ideal of measurable and null sets by $\Sigma_X$ and $\Null_X$, respectively. The morphisms of $\LMeas$ are the measurable maps $f : X \to Y$ modulo null sets, i.e.~$f,g : X\to Y$ represent the same morphism if $\{ x\in X \: | \: f(x) \neq g(x) \} \in \Null_X$.

A \emph{measure} $\mu$ on localizable measurable space $X$ is a measure $\mu : \Sigma_X \to \Rplusext$ in the usual sense, but such that every null set has zero measure. It hence descends to a map $\Sigma_X / \Null_X \to \Rplusext$ satisfying the inclusion-exclusion relation.

\tob{Is this map $\Sigma_X/\Null_X \to \Rplusext$ necessarily Scott continuous, so that it becomes a continuous valuation?}

>>>>

% categories
\newcommand{\cat}[1]{\mathsf{#1}}		% generic category
\newcommand{\Set}{\mathsf{Set}}
\newcommand{\Vect}[1]{\mathsf{Vect}_{#1}}	% vector spaces over a field
\newcommand{\Ban}{\mathsf{Ban}}			% (complex) Banach spaces
\newcommand{\Hilb}{\mathsf{Hilb}} 		% Hilb, cat of Hilbert spaces
\newcommand{\Rep}[1]{\mathsf{Rep}(#1)}		% cat normal Hilbert space reps of a C$^*$-algebra
\newcommand{\NRep}[1]{\mathsf{nRep}(#1)}	% cat of normal Hilbert space reps of a W$^*$-algebra
\newcommand{\id}[1]{1_{#1}}			% identity morphism
\newcommand{\Kay}{\mathcal{K}} % I use a convention that the name of a letter spelt phonetically signifies its mathcal version. But semantically, this is a compact class in a measure space.
\newcommand{\powerset}{\mathcal{P}}

@@

We let $\LMeas$ be the category of localizable measurable spaces~\cite{nlab:localizable_measure}. So objects of $\LMeas$ are triples $(X,\Sigma,\Null)$, where $\Sigma$ is a $\sigma$-algebra on $X$, and $\Null\subseteq\Sigma$ is a $\sigma$-ideal, whose elements are called \emph{null sets}, such that $\Sigma/\Null$ is complete as a Boolean algebra, and there exists a semifinite \cite[211F]{fremlin2} measure $\mu : \Sigma \rightarrow [0,\infty]$ such that $\Null = \mu^{-1}(0)$. These are what we get if we take a localizable measure space $(X,\Sigma,\mu)$  \cite[211G]{fremlin2} and only keep the information about which sets are null. Localizable measure spaces were introduced by Segal \cite{segal} as a characterization of the measure spaces such that $L^\infty(X)$, the C$^*$ of essentially bounded measurable functions modulo nullsets, is a W$^*$-algebra.

Measurable sets with null complement are also called \emph{full sets}. We usually just write $X$ for a localizable measurable space, and indicate its $\sigma$-algebra and $\sigma$-ideal of measurable and null sets by $\Sigma_X$ and $\Null_X$, respectively. The morphisms of $\LMeas$ are equivalence classes of measurable maps $f : X \to Y$ such that for all $N \in \Null_Y$, $f^{-1}(N) \in \Null_X$\todo[inline]{We need to be using compact strictly localizable spaces to make this definition sufficient to define a normal map between W*-algebras.}, under the following equivalence relation: $f \sim g$ iff for all $T \in \Sigma_Y$ $f^{-1}(T) \triangle g^{-1}(T) \in \Null_X$. If $f$ and $g$ agree almost everywhere, then $f \sim g$, and the two notions are equivalent if $(Y,\Sigma_Y)$ is countably separated \cite[343F]{fremlin3}, but not in general \cite[343I]{fremlin3}.

A \emph{measure} $\mu$ on localizable measurable space $X$ is a measure $\mu : \Sigma_X \to \Rplusext$ in the usual sense, but such that every null set has zero measure. It therefore descends to a countably additive map $\Sigma_X / \Null_X \to \Rplusext$, such that every non-zero element of $\Sigma_X / \Null_X$ has non-zero measure. As $\Sigma_X / \Null_X$ is a complete Boolean algebra, such a map is necessarily completely additive, not just countably additive \cite[321E]{fremlin2}.

Unfortunately, there exist localizable measurable spaces $X$, $Y$ such that $L^\infty(X) \cong L^\infty(Y)$ but there is no measurable map from $X$ to $Y$. To get out of this difficulty, we restrict ourselves to localizable measurable spaces arising from complete strictly localizable compact measure spaces.

To define these notions, we first define the finite intersection property, and what a compact class is. Let $X$ be a set, and $\Kay \subseteq \powerset(X)$ be a family of sets. We say it has the \emph{finite intersection property} if for each finite subfamily $(K_i)_{i \in I}$, $K_i \in \Kay$, we have $\bigcap_{i \in I} K_i \neq 0$. We say that $\Kay$ is a \emph{compact class} if for any subfamily $\Kay' \subseteq \Kay$ having the finite intersection property, $\bigcap \Kay' \neq \emptyset$. By taking the complement of every set in the open cover definition of a compact space, the reader may recognize that the closed sets of any compact topological space form a compact class. Slightly less trivially, the compact subsets of any Hausdorff space form a compact class.

A measure space $(X,\Sigma,\mu)$ is \emph{compact} \cite[342A (c)]{fremlin3} if there exists a compact class $\Kay$ such that for all $S \in \Sigma$
\[
\mu(S) = \sup \{ \mu(K) \mid K \in \Kay \cap \Sigma \text{ and } K \subseteq S \}.
\]
This condition is known as being \emph{inner regular with respect to} $\Kay$. Therefore any Radon measure on a Hausdorff space $X$ is compact, because we can take $\Kay$ to be the set of (topologically) compact subsets of $X$. So familiar examples, such as Lebesgue measure on $\R^n$, the Haar measure on a locally compact group, or any $\sigma$-finite measure on a standard Borel space, are compact.
